Cluster Analysis: Identifying Parkinson's Disease Subtypes

### 2016-03-03 writeup/abbreviated/abbreviated.tex
Almost done with second version of abbrv

\documentclass[10pt]{article}
\usepackage{fancyhdr}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{numprint}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
% Random packages from
% http://tex.stackexchange.com/questions/50070/landscape-figure-in-latex
% Necessary for sideways pictures
\usepackage{wrapfig}
\usepackage{lscape}
\usepackage{rotating}
\usepackage{epstopdf}
\usepackage{tablefootnote}
% for word wrap verbatim
\usepackage{listings}
\usepackage{caption}
\usepackage{subfig}
\lstset{
   breaklines=true,
   basicstyle=\ttfamily}
% \pagestyle{fancy}
% \lhead{Jesse Mu}
% \rhead{CSCI339 Term Project}
% One more .. since we're in the abbreviated subdir
\graphicspath{ {../../figures/} }
\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}}

@@

@@

\begin{table}[h]
  \centering
  \begin{tabular}{l|l|l}
    Name & Type & Description \\
    \hline
    nms\_d1 & byte & cardiovascular \\
    nms\_d2 & byte & sleep/fatigue \\
    nms\_d3 & byte & mood/cognition \\
    nms\_d4 & byte & percep/hallucinations \\
    nms\_d5 & byte & attention/memory \\
    nms\_d6 & byte & gastrointestinal \\
    nms\_d7 & byte & urinary \\
    nms\_d8 & byte & sexual function \\
    nms\_d9 & byte & miscellaneous \\
    tremor & float & tremor \\
    bradykin & float & bradykinesia\tablefootnote{Impaired ability to
    adjust the body's position.} \\
    rigidity & float & rigidity \\
    axial & float & axial\tablefootnote{Issues affecting the middle of
    the body.} \\
  \end{tabular}
  \caption{Selected Features and Details}
  \label{tab:selected-features}
\end{table}

\begin{table}[h]
  \centering
  \begin{tabular}{l|l|l|l}
  Name  &       $\mu$ & $\sigma$ & min-max \\
         \hline
nms\_d1&   1.73&  3.35&   0-24 \\
nms\_d2&   8.75&  8.70&   0-48 \\
nms\_d3&   8.68& 11.55&   0-60 \\
nms\_d4&   1.64&  3.86&   0-33 \\
nms\_d5&   5.42&  7.43&   0-36 \\
nms\_d6&   5.53&  6.79&   0-36 \\
nms\_d7&   8.08&  8.94&   0-36 \\
nms\_d8&   3.52&  5.97&   0-24 \\
nms\_d9&   7.13&  7.79&   0-48 \\
tremor&   2.59&  2.58&   0-12 \\
bradykin& 2.40&  1.41&   0-6 \\
rigidity& 2.24&  1.36&   0-6 \\
axial&    3.25&  2.68&   0-12 \\
  \end{tabular}
  \caption{Descriptive Statistics}
  \label{tab:descriptive-statistics}
\end{table}

@@

$k$-means clustering with $k = 4$ was tried. Statistics for determining the
optimal number of clusters were used, but were inconclusive: results in
Figure~\ref{tab:numclus}. This probably indicates that the data is not very
well clustered. $k = 2, 3$ provided models that were too simplistic. $k = 5$
did not provide any new information, but rather just fragmented existing
groups.
\begin{table}[h]
  \centering
  \begin{tabular}{l|l}
    Criterion & Optimal $k$ \\
    \hline
    Minimum ASW & 2 \\
    BIC & 18 \\
    SSE Scree Plot & Inconclusive \\
    Gap Statistic & 4 \\
    Affinity Propagation & 8 \\%\tablefootnote{$\lambda = 0.98$, q = 0, maxits = 1000, %convits = 100} & 8 \\

  \end{tabular}
  \caption{Results of various techniques for determining $k$}
  \label{tab:numclus}
\end{table}

\subsection{Decision tree}
% \begin{table}[h]
  % \centering
  % \begin{tabular}{l|l|l|l|l|l}
    % $k$ & CP\tablefootnote{Complexity Parameter} & CV Xerror\tablefootnote{10-fold cross
    % validation} & Root Feature &
    % Root Error & Figure \\
    % \hline
    % 4 & 0.0100 & 0.255 & pigd $<$ 2.5 & 0.563 & Figure~\ref{fig:kmeans-dtree-4} \\
  % \end{tabular}
  % \caption{$k$-kmeans decision trees statistics}
  % \label{tab:k-means-dtrees}
% \end{table}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{dtree-kmeans-pruned-unscaled-4.pdf}
  \caption{Decision Tree from $k$-means clustering, 4 clusters}
  \label{fig:kmeans-dtree-4}
\end{figure}

@@

\begin{figure*}[h]
  \centering
  \includegraphics[width=\linewidth]{kmeans-summaries-4.pdf}
  \caption{Cluster Summaries, $k = 4$}
  \label{fig:kmeans-summaries-4}
\end{figure*}

@@

\begin{table}[h]
  \centering
  \begin{tabular}{l|l|l}
  Variable & Cluster Relation & $p$ \\
  \hline
age & 2-1 & 0.428 \\
    & 3-2 & 0.724 \\
    \hline
sex & 2-1 & 0.0918 \\
    & 3-1 & 0.216 \\
    & 4-1 & 0.827 \\
    & 4-2 & 0.849 \\
    & 4-3 & 0.161 \\
    \hline
pdonset & 2-1 & 0.859 \\
        & 3-1 & 0.700 \\
        & 4-1 & 0.305 \\
        & 3-2 & 0.370 \\
        & 4-2 & 0.147 \\
        & 4-3 & 0.803 \\
\hline
durat\_pd & 3-2 & 0.562 \\
\hline
cisitot & 3-2 & 0.522 \\
\hline
nms\_d1 & 3-1 & 0.333 \\
\hline
nms\_d4 & 3-1 & 0.557 \\
\hline
nms\_d5 & 3-1 & 0.856 \\
\hline
nms\_d8 & 3-1 & 0.122 \\
\hline
nms\_d9 & 3-1 & 0.0735 \\
       & 4-2 & 0.730 \\
\hline
tremor & 2-1 & 0.360 \\
  \end{tabular}
  \caption{Tukey's HSD Insignificant Differences}
  \label{tab:tukeyhsd}
\end{table}

@@

\begin{table}[h]
  \centering
  \begin{tabular}{r|l|l}
    rank & variable & information gain \\
    \hline
    1 & bradykin      & 0.316 \\
    2 & rigidity      & 0.296 \\
    3 & nms\_d2      & 0.242 \\
    4 & cisitot      & 0.229 \\
    5 & axial      & 0.228 \\
    6 & nms\_d3      & 0.205 \\
    7 & nms\_d9      & 0.158 \\
    8 & nms\_d7      & 0.153 \\
    9 & nms\_d5      & 0.145 \\
    10 & nms\_d6      & 0.140 \\
    11 & nms\_d1      & 0.132 \\
    12 & tremor      & 0.109 \\
    13 & nms\_d4      & 0.107 \\
    14 & nms\_d8      & 0.100 \\
    15 & durat\_pd      & 0.0288 \\
    16 & age      & 0.0235 \\
    17 & sex      & 0.000 \\
    18 & pdonset      & 0.000 \\
  \end{tabular}
  \caption{Features ranked by information gain}
  \label{tab:info_gain}
\end{table}

@@

\begin{table*}[h]
  \centering
  \begin{tabular}{l|l|l|l}
 Cluster & Variables & 95\% CI & $p$ \\
    \hline
1 & bradykin, cisitot&[-0.0225, 0.171]&0.131 \\
  & rigidity, cisitot&[-0.000406, 0.192]&0.0510 \\
  & bradykin, nms\_d6&[-0.0634, 0.131]&0.493 \\
  & rigidity, nms\_d6&[-0.101, 0.0932]&0.934 \\
  \hline
2 & bradykin, cisitot&[0.0786, 0.351]&$0.00248^{(**)}$ \\
  & rigidity, cisitot&[-0.215, 0.069]&0.310 \\
  & bradykin, nms\_d6&[-0.152, 0.133]&0.897 \\
  & rigidity, nms\_d6&[-0.123, 0.163]&0.781 \\
  \hline
3 & bradykin, cisitot&[0.0995, 0.350]&$0.000620^{(***)}$ \\
  & rigidity, cisitot&[0.0687, 0.322]&$0.00298^{(**)}$ \\
  & bradykin, nms\_d6&[0.0350, 0.292]&$0.0134^{(*)}$ \\
  & rigidity, nms\_d6&[-0.0846, 0.179]&0.478 \\
  \hline
  4 & bradykin, cisitot&[0.454, 0.724]&${3.97\e{-10}}^{(***)}$ \\
& rigidity, cisitot&[0.375, 0.675]&${4.99\e{-08}}^{(***)}$ \\
  & bradykin, nms\_d6&[0.297, 0.624]&${2.60\e{-06}}^{(***)}$ \\
  & rigidity, nms\_d6&[0.278, 0.611]&${6.43\e{-06}}^{(***)}$ \\
  \end{tabular}
  \caption{Correlation tests. (*) $p < 0.05$, (**) $p < 0.01$, (***), $p < 0.001$}
  \label{tab:cortests}
\end{table*}

\begin{figure}[h]
  \centering
  \subfloat[][]{%
  \includegraphics[clip,width=0.8\linewidth]{bradykin-v-cisitot.pdf}
  \label{bradyv}
}

\subfloat[][]{%
  \includegraphics[clip,width=0.8\linewidth]{rigidity-v-cisitot.pdf}
  \label{rigidv}
}
  \caption{Relationship between \protect\subref{bradyv} bradykinesia,
  \protect\subref{rigidv} rigidity and overall severity (cisitot). Shaded band
is 95\% confidence interval.}
  \label{fig:corrxy-br}
\end{figure}

\begin{figure}[h]
  \centering
  \subfloat[][]{%
  \includegraphics[clip,width=0.8\linewidth]{bradykin-v-nms_d6.pdf}
  \label{brady6}
}

\subfloat[][]{%
  \includegraphics[clip,width=0.8\linewidth]{rigidity-v-nms_d6.pdf}
  \label{rigid6}
}
  \caption{Relationship between \protect\subref{brady6} bradykinesia,
  \protect\subref{rigid6} rigidity and nms\_d6 (gastrointestinal). Shaded band
is 95\% confidence interval.}
  \label{fig:corrxy-d6}
\end{figure}
\begin{figure}[h]
  \centering
  \includegraphics[width=0.68\linewidth]{corrplots.pdf}
  \caption{Correlation plots}
  \label{fig:corrplots}
\end{figure}

@@

\begin{table}[h]
  \centering
  \begin{tabular}{l|l}
    Criterion & Optimal $k$ \\
    \hline
    Minimum ASW & 2 \\
    BIC & 1 (?) \\
    SSE Scree Plot & Inconclusive \\
    Gap Statistic & 3 \\
    Affinity Propagation\tablefootnote{$\lambda = 0.98$, q = 0, maxits = 1000,
    convits = 100} & 5 \\
  \end{tabular}
  \caption{Results of various techniques for determining $k$, applied to
  subtype 2}
  \label{tab:numclus-nms}
\end{table}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{c2-summaries-2.pdf}
  \caption{Clustering on nonmotor group: $k = 2$}
  \label{fig:c2-summaries-2}
\end{figure}
\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{c2-summaries-3.pdf}
  \caption{Clustering on nonmotor group: $k = 3$}
  \label{fig:c2-summaries-3}
\end{figure}
\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{c2-summaries-4.pdf}
  \caption{Clustering on nonmotor group: $k = 4$}
  \label{fig:c2-summaries-4}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=0.6\linewidth]{dtree-1va-pruned.pdf}
  \caption{Cluster 1 (mild) vs all}
  \label{fig:1va}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=0.6\linewidth]{dtree-2va-pruned.pdf}
  \caption{Cluster 2 (nonmotor-dominated) vs all}
  \label{fig:2va}
\end{figure}
\begin{figure}[h]
  \centering
  \includegraphics[width=0.6\linewidth]{dtree-3va-pruned.pdf}
  \caption{Cluster 3 (motor-dominated) vs all}
  \label{fig:3va}
\end{figure}
\begin{figure}[h]
  \centering
  \includegraphics[width=0.6\linewidth]{dtree-4va-pruned.pdf}
  \caption{Cluster 4 (severe) vs all}
  \label{fig:4va}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=0.6\linewidth]{dtree-2and4va-pruned.pdf}
  \caption{Clusters 2 (nms) and 4 (severe) vs rest (1 and 3)}
  \label{fig:dtree-2and4va-pruned}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=0.6\linewidth]{dtree-2and3va-pruned.pdf}
  \caption{Clusters 2 (nms) and 3 (motor) vs rest (1 and 4)}
  \label{fig:dtree-2and3va-pruned}
\end{figure}

@@

>>>>

\documentclass[10pt]{article}
\usepackage{fancyhdr}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{numprint}
\usepackage{float}
\usepackage[margin=0.75in]{geometry}
\usepackage{graphicx}
% Random packages from
% http://tex.stackexchange.com/questions/50070/landscape-figure-in-latex
% Necessary for sideways pictures
\usepackage{wrapfig}
\usepackage{lscape}
\usepackage{rotating}
\usepackage{epstopdf}
\usepackage{tablefootnote}
% for word wrap verbatim
\usepackage{listings}
\usepackage{caption}
\usepackage{multicol}
\usepackage{subfig}
\lstset{
   breaklines=true,
   basicstyle=\ttfamily}
% Include longitudinal
\graphicspath{ {../../figures/} {../../figures/longitudinal/} }
\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}}

@@

\tableofcontents
\newpage

\begin{multicols}{2}

\section{New Content}

Section~\ref{sec:longitudinal} on longitudinal analysis; section~\ref{sec:nms30} on

@@

\begin{table}[H]
  \centering
  \begin{tabular}{l|l|l}
    Name & Type & Description \\
    \hline
    nms\_d1 & byte & cardiovascular \\
    nms\_d2 & byte & sleep/fatigue \\
    nms\_d3 & byte & mood/cognition \\
    nms\_d4 & byte & percep/hallucinations \\
    nms\_d5 & byte & attention/memory \\
    nms\_d6 & byte & gastrointestinal \\
    nms\_d7 & byte & urinary \\
    nms\_d8 & byte & sexual function \\
    nms\_d9 & byte & miscellaneous \\
    tremor & float & tremor \\
    bradykin & float & bradykinesia\tablefootnote{Impaired ability to
    adjust the body's position.} \\
    rigidity & float & rigidity \\
    axial & float & axial\tablefootnote{Issues affecting the middle of
    the body.} \\
  \end{tabular}
  \caption{Selected Features and Details}
  \label{tab:selected-features}
\end{table}

\begin{table}[H]
  \centering
  \begin{tabular}{l|l|l|l}
  Name  &       $\mu$ & $\sigma$ & min-max \\
         \hline
nms\_d1&   1.73&  3.35&   0-24 \\
nms\_d2&   8.75&  8.70&   0-48 \\
nms\_d3&   8.68& 11.55&   0-60 \\
nms\_d4&   1.64&  3.86&   0-33 \\
nms\_d5&   5.42&  7.43&   0-36 \\
nms\_d6&   5.53&  6.79&   0-36 \\
nms\_d7&   8.08&  8.94&   0-36 \\
nms\_d8&   3.52&  5.97&   0-24 \\
nms\_d9&   7.13&  7.79&   0-48 \\
tremor&   2.59&  2.58&   0-12 \\
bradykin& 2.40&  1.41&   0-6 \\
rigidity& 2.24&  1.36&   0-6 \\
axial&    3.25&  2.68&   0-12 \\
  \end{tabular}
  \caption{Descriptive Statistics}
  \label{tab:descriptive-statistics}
\end{table}

@@

$k$-means clustering with $k = 4$ was tried. Statistics for determining the
optimal number of clusters were used, but were generally inconclusive: results in
Figure~\ref{tab:numclus}. This probably indicates that the data is not very
well clustered. $k = 2, 3$ provided models that were too simplistic. $k = 5$
did not provide any new information, but rather just fragmented existing
groups.
\begin{table}[H]
  \centering
  \begin{tabular}{l|l}
    Criterion & Optimal $k$ \\
    \hline
    Minimum ASW & 2 \\
    BIC & 18 \\
    SSE Scree Plot & Inconclusive \\
    Gap Statistic & 4 \\
    Affinity Propagation & 8 \\%\tablefootnote{$\lambda = 0.98$, q = 0, maxits = 1000, %convits = 100} & 8 \\
    clValid stability measures & 4 \\
  \end{tabular}
  \caption{Results of various techniques for determining $k$}
  \label{tab:numclus}
\end{table}

\subsection{Decision tree}
% \begin{table}[H]
  % \centering
  % \begin{tabular}{l|l|l|l|l|l}
    % $k$ & CP\tablefootnote{Complexity Parameter} & CV Xerror\tablefootnote{10-fold cross
    % validation} & Root Feature &
    % Root Error & Figure \\
    % \hline
    % 4 & 0.0100 & 0.255 & pigd $<$ 2.5 & 0.563 & Figure~\ref{fig:kmeans-dtree-4} \\
  % \end{tabular}
  % \caption{$k$-kmeans decision trees statistics}
  % \label{tab:k-means-dtrees}
% \end{table}

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=\linewidth]{dtree-kmeans-pruned-unscaled-4.pdf}
  \caption{Decision Tree from $k$-means clustering, 4 clusters}
  \label{fig:kmeans-dtree-4}
\end{figure}

@@

\begin{figure*}[t]
  \centering
  \includegraphics[width=\linewidth]{kmeans-summaries-4.pdf}
  \caption{Cluster Summaries, $k = 4$}
  \label{fig:kmeans-summaries-4}
\end{figure*}

@@

\begin{table}[H]
  \centering
  \begin{tabular}{l|l|l}
  Variable & Cluster Relation & $p$ \\
  \hline
age & 2-1 & 0.428 \\
    & 3-2 & 0.724 \\
    \hline
sex & 2-1 & 0.0918 \\
    & 3-1 & 0.216 \\
    & 4-1 & 0.827 \\
    & 4-2 & 0.849 \\
    & 4-3 & 0.161 \\
    \hline
pdonset & 2-1 & 0.859 \\
        & 3-1 & 0.700 \\
        & 4-1 & 0.305 \\
        & 3-2 & 0.370 \\
        & 4-2 & 0.147 \\
        & 4-3 & 0.803 \\
\hline
durat\_pd & 3-2 & 0.562 \\
\hline
cisitot & 3-2 & 0.522 \\
\hline
nms\_d1 & 3-1 & 0.333 \\
\hline
nms\_d4 & 3-1 & 0.557 \\
\hline
nms\_d5 & 3-1 & 0.856 \\
\hline
nms\_d8 & 3-1 & 0.122 \\
\hline
nms\_d9 & 3-1 & 0.0735 \\
       & 4-2 & 0.730 \\
\hline
tremor & 2-1 & 0.360 \\
  \end{tabular}
  \caption{Tukey's HSD Insignificant Differences}
  \label{tab:tukeyhsd}
\end{table}

@@

\begin{table}[H]
  \centering
  \begin{tabular}{r|l|l}
    rank & variable & information gain \\
    \hline
    1 & bradykin      & 0.316 \\
    2 & rigidity      & 0.296 \\
    3 & nms\_d2      & 0.242 \\
    4 & cisitot      & 0.229 \\
    5 & axial      & 0.228 \\
    6 & nms\_d3      & 0.205 \\
    7 & nms\_d9      & 0.158 \\
    8 & nms\_d7      & 0.153 \\
    9 & nms\_d5      & 0.145 \\
    10 & nms\_d6      & 0.140 \\
    11 & nms\_d1      & 0.132 \\
    12 & tremor      & 0.109 \\
    13 & nms\_d4      & 0.107 \\
    14 & nms\_d8      & 0.100 \\
    15 & durat\_pd      & 0.0288 \\
    16 & age      & 0.0235 \\
    17 & sex      & 0.000 \\
    18 & pdonset      & 0.000 \\
  \end{tabular}
  \caption{Features ranked by information gain}
  \label{tab:info_gain}
\end{table}

@@

\begin{table*}[t]
  \centering
  \begin{tabular}{l|l|l|l}
 Cluster & Variables & 95\% CI & $p$ \\
    \hline
1 & bradykin, cisitot&[-0.0225, 0.171]&0.131 \\
  & rigidity, cisitot&[-0.000406, 0.192]&0.0510 \\
  & bradykin, nms\_d6&[-0.0634, 0.131]&0.493 \\
  & rigidity, nms\_d6&[-0.101, 0.0932]&0.934 \\
  \hline
2 & bradykin, cisitot&[0.0786, 0.351]&$0.00248^{(**)}$ \\
  & rigidity, cisitot&[-0.215, 0.069]&0.310 \\
  & bradykin, nms\_d6&[-0.152, 0.133]&0.897 \\
  & rigidity, nms\_d6&[-0.123, 0.163]&0.781 \\
  \hline
3 & bradykin, cisitot&[0.0995, 0.350]&$0.000620^{(***)}$ \\
  & rigidity, cisitot&[0.0687, 0.322]&$0.00298^{(**)}$ \\
  & bradykin, nms\_d6&[0.0350, 0.292]&$0.0134^{(*)}$ \\
  & rigidity, nms\_d6&[-0.0846, 0.179]&0.478 \\
  \hline
  4 & bradykin, cisitot&[0.454, 0.724]&${3.97\e{-10}}^{(***)}$ \\
& rigidity, cisitot&[0.375, 0.675]&${4.99\e{-08}}^{(***)}$ \\
  & bradykin, nms\_d6&[0.297, 0.624]&${2.60\e{-06}}^{(***)}$ \\
  & rigidity, nms\_d6&[0.278, 0.611]&${6.43\e{-06}}^{(***)}$ \\
  \end{tabular}
  \caption{Correlation tests. (*) $p < 0.05$, (**) $p < 0.01$, (***), $p < 0.001$}
  \label{tab:cortests}
\end{table*}

\begin{figure}[H]
  \centering
  \subfloat[][]{%
  \includegraphics[clip,width=\linewidth]{bradykin-v-cisitot.pdf}
  \label{bradyv}
}

\subfloat[][]{%
  \includegraphics[clip,width=\linewidth]{rigidity-v-cisitot.pdf}
  \label{rigidv}
}
  \caption{Relationship between \protect\subref{bradyv} bradykinesia,
  \protect\subref{rigidv} rigidity and overall severity (cisitot). Shaded band
is 95\% confidence interval.}
  \label{fig:corrxy-br}
\end{figure}

\begin{figure}[H]
  \centering
  \subfloat[][]{%
  \includegraphics[clip,width=\linewidth]{bradykin-v-nms_d6.pdf}
  \label{brady6}
}

\subfloat[][]{%
  \includegraphics[clip,width=\linewidth]{rigidity-v-nms_d6.pdf}
  \label{rigid6}
}
  \caption{Relationship between \protect\subref{brady6} bradykinesia,
  \protect\subref{rigid6} rigidity and nms\_d6 (gastrointestinal). Shaded band
is 95\% confidence interval.}
  \label{fig:corrxy-d6}
\end{figure}
\begin{figure}[H]
  \centering
  \includegraphics[width=\linewidth]{corrplots.pdf}
  \caption{Correlation plots}
  \label{fig:corrplots}
\end{figure}

@@

\begin{table}[H]
  \centering
  \begin{tabular}{l|l}
    Criterion & Optimal $k$ \\
    \hline
    Minimum ASW & 2 \\
    BIC & 1 (?) \\
    SSE Scree Plot & Inconclusive \\
    Gap Statistic & 3 \\
    Affinity Propagation\tablefootnote{$\lambda = 0.98$, q = 0, maxits = 1000,
    convits = 100} & 5 \\
  \end{tabular}
  \caption{Results of various techniques for determining $k$, applied to
  subtype 2}
  \label{tab:numclus-nms}
\end{table}

@@

\begin{figure*}[p]
  \centering
  \includegraphics[width=0.65\linewidth]{c2-summaries-2.pdf}
  \caption{Clustering on nonmotor group: $k = 2$}
  \label{fig:c2-summaries-2}
\end{figure*}
\begin{figure*}[p]
  \centering
  \includegraphics[width=0.65\linewidth]{c2-summaries-3.pdf}
  \caption{Clustering on nonmotor group: $k = 3$}
  \label{fig:c2-summaries-3}
\end{figure*}
\begin{figure*}[p]
  \centering
  \includegraphics[width=0.65\linewidth]{c2-summaries-4.pdf}
  \caption{Clustering on nonmotor group: $k = 4$}
  \label{fig:c2-summaries-4}
\end{figure*}

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=0.75\linewidth]{dtree-1va-pruned.pdf}
  \caption{Cluster 1 (mild) vs all}
  \label{fig:1va}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=0.75\linewidth]{dtree-2va-pruned.pdf}
  \caption{Cluster 2 (nonmotor-dominated) vs all}
  \label{fig:2va}
\end{figure}
\begin{figure}[H]
  \centering
  \includegraphics[width=0.75\linewidth]{dtree-3va-pruned.pdf}
  \caption{Cluster 3 (motor-dominated) vs all}
  \label{fig:3va}
\end{figure}
\begin{figure}[H]
  \centering
  \includegraphics[width=0.75\linewidth]{dtree-4va-pruned.pdf}
  \caption{Cluster 4 (severe) vs all}
  \label{fig:4va}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=0.75\linewidth]{dtree-2and4va-pruned.pdf}
  \caption{Clusters 2 (nms) and 4 (severe) vs rest (1 and 3)}
  \label{fig:dtree-2and4va-pruned}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=0.75\linewidth]{dtree-2and3va-pruned.pdf}
  \caption{Clusters 2 (nms) and 3 (motor) vs rest (1 and 4)}
  \label{fig:dtree-2and3va-pruned}
\end{figure}

@@

\section{Longitudinal Analysis}
\label{sec:longitudinal}

I wanted to observe the progression of nonmotor symptoms of Parkinson's disease akin to work done
by various studies (e.g.\ \cite{onsetpd,vu12,zahodne12}). The data was a little bit too noisy when
I tried to examine correlations, so I decided to average the values into bins for
\texttt{durat\_pd}: 0-2 years, 2-4 years, 4-6 years, etc. First, the correlation based on PD
duration for each symptom is displayed in Figure~\ref{fig:longcorr}.

\begin{quote}
  Note: I have not included all of the graphs I have, just those I found interesting. If you would
  like all of the graphs, I can send those to you in a separate email.
\end{quote}

\begin{figure*}[t]
  \centering
  \includegraphics[width=0.9\linewidth]{pd-durat-cor.pdf}
  \caption{Correlation of variables according to PD Duration}
  \label{fig:longcorr}
\end{figure*}

Most symptoms have at least a minor positive correlation. Interestingly, tremor, nms9 (anxiety)
and nms10 (depression) do not have a strong correlation with PD duration.
% This could give evidence that whether or not someone has high tremor is a distinct subtype?

\subsection{Anxiety}
I plotted the mean nms\_9 (anxiety) score for each durat\_pd bin in
Figures~\ref{fig:nms9-long} and ~\ref{fig:nms9-multi}

\begin{figure*}[p]
  \centering
  \includegraphics[width=0.8\linewidth]{nms9-durat-counts.pdf}
  \caption{Mean anxiety (nms\_9) score by PD duration, with number of patients in each group at
  bottom.}
  \label{fig:nms9-long}
\end{figure*}

\begin{figure*}[p]
  \centering
  \includegraphics[width=0.8\linewidth]{nms9-multi-durat-counts.pdf}
  \caption{Mean anxiety (nms\_9) score per subtype, with number of patients in each group at
  bottom.}
  \label{fig:nms9-multi}
\end{figure*}

\subsection{Depression}
I plotted the mean nms\_10 (depression) score for each durat\_pd bin in
Figures~\ref{fig:nms10-long} and ~\ref{fig:nms10-multi}.

\begin{figure*}[p]
  \centering
  \includegraphics[width=0.8\linewidth]{nms10-durat-counts.pdf}
  \caption{Mean depression (nms\_10) score by PD duration, with number of patients in each group at
  bottom.}
  \label{fig:nms10-long}
\end{figure*}

\begin{figure*}[p]
  \centering
  \includegraphics[width=0.8\linewidth]{nms10-multi-durat-counts.pdf}
  \caption{Mean depression (nms\_10) score per subtype, with number of patients in each group at
  bottom.}
  \label{fig:nms10-multi}
\end{figure*}

\subsection{Cisitot}

Note that cisitot has the highest correlation of all of the PD symptoms. I plotted cisitot per
subtype in Figure~\ref{fig:cisitot-multi}. Interestingly, subtypes 2, 3, and 4 generally start from
the same mean motor symptom score, suggesting different paths of disease progression.
% same global mean motor symptom score. % TODO: This is important.

\begin{figure*}[t]
  \centering
  \includegraphics[width=0.8\linewidth]{cisitot-multi-durat-counts.pdf}
  \caption{Mean cisitot score per subtype, with number of patients in each group at
  bottom.}
  \label{fig:cisitot-multi}
\end{figure*}

\subsection{Tremor}

I plotted tremor symptoms in Figure~\ref{fig:tremor-multi}. Notice how in this group, subtype 3
(motor-dom) is higher than all groups towards the beginning of disease progression. This supports
the idea that Subtype 3 is Ma's \cite{ma15} tremor-dominant cluster.

\begin{figure*}[t]
  \centering
  \includegraphics[width=0.8\linewidth]{tremor-multi-durat-counts.pdf}
  \caption{Mean tremor score per subtype, with number of patients in each group at
  bottom.}
  \label{fig:tremor-multi}
\end{figure*}

% TODO: Maybe later
% \subsection{By \texttt{pdonset}}

\section{Extended Nonmotor Symptoms}
\label{sec:nms30}

\subsection{Adding nonmotor symptoms to original $k$-means clustering}

First, I used the clustering with the original 9 nms symptoms, but plotted the original 30 symptoms
instead. The reason for this was to treat the clustering of the 9 symptoms as a form of
dimensionality reduction, since we know each symptom of nms\{1-30\} is colinear with several other
symptoms, and they form the broader nms\_d\{1-9\} symptoms. The results are in
Figure~\ref{fig:d9to30}

\begin{figure*}[p]
  \centering
  \includegraphics[width=\linewidth]{d9to30.png}
  \caption{All symptoms (nms\{1-30\}) plotted with the original clustering results of the the 9
broader symptoms. Interesting parts of the figure are are labeled in red.}
  \label{fig:d9to30}
\end{figure*}

\subsection{\emph{New} clustering with 30 symptoms}

\subsubsection{Principal Component Analysis}

I ran PCA on the dataset with solely the 30 nonmotor symptoms. A plot of various metrics for
determining the optimal number of eigenvectors is in Figure~\ref{fig:nms30only-eigs}. Most metrics agree
Nthat the optimal number of eigenvalues is around 6, but notice how one component has by far the
highest eigenvalue; this component can be interpreted as a general ``nonmotor severity'' measure,
indicating that much of the general nonmotor symptomatic expression falls along the same dimension.

\begin{figure}[H]
  \centering
  \includegraphics[width=\linewidth]{nms30only-eigs.pdf}
  \caption{Eigenvalues and other metrics for Principal Component Analysis applied to the NMS30 data.}
  \label{fig:nms30only-eigs}
\end{figure}

This principal component can be shown more clearly in the correlation plot of each symptom with the
first 5 components located in Figure~\ref{fig:nms30only-eigscorr}. In the first component, all of the
symptoms have at least a minor positive correlation. The remaining components are composed
primarily of dimensionality reduction in domain 3 (mood/cognition), domain 5 (attention/memory),
domain 7 (urinary), and domain 9 (misc). It is once again demonstrated that mood/cognition symptoms
are very important parts of this dataset.

\begin{figure*}[t]
  \centering
  \includegraphics[width=\linewidth]{nms30only-eigscorr.pdf}
  \caption{Correlation of each nonmotor symptom with the first 6 components}
  \label{fig:nms30only-eigscorr}
\end{figure*}

\subsubsection{Symptoms clustering}

\paragraph{Nonmotor only}

I used hierarchical clustering to cluster the 30 symptoms of the dataset using complete linkage.
Results are in Figure~\ref{fig:nms30only-colhc}. Naturally, symptoms that belong to the same domain are
clustered together. But the hierarchical clustering also shows how some symptomatic domains are
closely related; for example, d6 (gastro), d7 (urinary), and d8(sexual). Additionally,
miscellaneous symptoms are quite distinct from the other symptoms.

\begin{figure*}[p]
  \centering
  \includegraphics[width=\linewidth]{nms30only-colhc.pdf} \caption{Hierarchical clustering of symptoms.
    Dendrogram colored with 5 clusters. Naming is: d\{1-9\}-\{1-30\}-description}
  \label{fig:nms30only-colhc}
\end{figure*}

\paragraph{With motor symptoms}

In Figure~\ref{fig:nms30m-colhc}, the results are similar to the original hierarchical
clustering; but notice that motor symptoms are clustered together \emph{except} for tremor. In
fact, tremor is the most dissimilar symptom of them all!

\begin{figure*}[p]
  \centering
  \includegraphics[width=\linewidth]{nms30m-colhc.pdf} \caption{Hierarchical clustering of symptoms
  including motor symptoms.  Dendrogram colored with 4 clusters.}
  \label{fig:nms30m-colhc}
\end{figure*}

\subsubsection{$k$-means}

More so than previously, metrics for identifying the number of clusters proved inconclusive. Most
metrics suggested 2 clusters, which isn't a useful clustering. I decided to stick with $k = 4$ to
compare the results of this new clustering with the previous run. All 30 NMS symptoms were included
as well as the four motor symptoms axial, rigidity, bradykinesia, and tremor.
In this clustering, 4 slightly different subtypes are identified:

\begin{enumerate}
  \item ($n = 509$) Mildly affected in all domains.
  \item ($n = 97$) Higher than average nonmotor symptoms, but severely affected in depressive
    symptoms especially.
  \item ($n = 249$) Averagely affected in all domains.
  \item ($n = 49$) Severely affected in all domains.
\end{enumerate}

Because the number of symptoms is high, I don't provide the boxplots I have previously provided;
instead, I provide a heatmap of $z$-scores in Figure~\ref{fig:nms30m-k4-heatmap-improved} and a
table of results in Table~\ref{tab:nms30_k4}.

\begin{figure*}[p]
  \centering
  \includegraphics[width=0.6\linewidth]{nms30m-k4-heatmap-improved.png}
  \caption{Heatmap for $k$-means clustering, $k = 4$, with the 30 nonmotor symptoms. Notice that
  subtype 2 is a ``Depression-Dominant'' subtype.} \label{fig:nms30m-k4-heatmap-improved}
\end{figure*}
